Museum specimens as baseline for assessing a potential global pollinator crisis

### 2018-02-02 ms.Rmd

Building on previous studies showing the use of historical museum specimens to compare historical bee occurrences in northeast USA, Central Europe, and South Africa with current occurrence patterns, we show the potential for Museum specimens to play a key role in elucidating the extent of the pollinator crisis and compare the responses obtained along different environmental situations. Highlight knowledge gaps and oportunities (e.g. London museum, zooniverse...). Interestingly, we will add data from New Zealand and Spain increasing the geographical coverage and including bee diversity hotspots like Spain, for which bee declines have never been assessed.

@@

##Current evidence:
- US: Bartomeus 2013 and Burkle 201?, ... NAS book Critizize Burkle dessign and compare to mine.
- EU: Beijsmier, Carvalho, Schepper, ... UK? Peter Keil flies? Moths? Buterflies!
- South Africa: Pauw, maybe with plants...
- Specific examples e.g. Bombus dalbhomi (Aizen), ... Thomson ecology letters

@@

## Future needs /knowledge gaps:
- Digitalization of MORE museums (coordinated)
- Integration of DATA and analysis:
- Communication Museums- Sci (biderictional. Importance of voucering specimens fused for science).
- Compare global differences geographically (Eu Z US, for example)

@@

>>>>

Statements about pollinator crisis, or bee declines are widespread, but information is currenti limiting....
Building on previous studies showing the use of historical museum specimens to compare historical bee occurrences in northeast USA, Central Europe, and South Africa with current occurrence patterns, we show the potential for Museum specimens to play a key role in elucidating the extent of the pollinator crisis and compare the responses obtained along different environmental situations. Highlight knowledge gaps and oportunities (e.g. London museum, zooniverse...). Interestingly, we will add data from New Zealand and Spain increasing the geographical coverage and including bee diversity hotspots like Spain, for which bee declines have never been assessed.

@@

##Current evidence:
- US: Bartomeus 2013 and Burkle 201?, ... NAS book Critizize Burkle dessign and compare to mine.
- EU: Beijsmier, Carvalho, Schepper, ... UK? Peter Keil flies? Moths? Buterflies!
- South Africa: Pauw, maybe with plants...
- Specific examples e.g. Bombus dalbhomi (Aizen), ... Thomson ecology letters
 Fig 1: a global map showing evidence (% decline in richness) and Data deficient areas would be great.

@@

## Future needs /knowledge gaps:
- Digitalization of MORE museums (coordinated) Just common protocols? RopenSci? What about an interactive bee decline map? Predicts?
- Integration of DATA and analysis:
- Communication Museums- Sci (biderictional. Importance of voucering specimens fused for science).
- Compare global differences geographically (Eu Z US, for example)

@@

#Acknowledgements

E. Asensio (or maybe coauthor?). Patrick McQuinn, and Crona McMonagle for data entry. Madrid museum, ITACyL.
